---
title: "Hockey Project"
author: "Jackie Steffan"
date: "9/11/2020"
output: 
    rmarkdown ::github_document:
      toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(devtools)
library(tidyverse)
library(httr)
library(jsonlite)
library(ggplot2)
library(DT)
```
```{r eval=FALSE}
render("hockey.md", output_file = "README.md")
```

Github Repo: [https://github.com/JackieSteffan/Hockey](https://github.com/JackieSteffan/Hockey)  

# Required Packages

Load in the following packages:  
`rmarkdown`, `tidyverse`, `devtools`, `httr`, `jsonlite`, `ggplot2`, `DT`

# Create Functions

```{r functions, warning= FALSE}
franchise <- function(team, type) {
  franchise_url <- paste0("https://records.nhl.com/site/api/franchise")
  get_franchise <- GET(franchise_url)
  txt_franchise <- content(get_franchise, "text")
  json_franchise <- fromJSON(txt_franchise, flatten=TRUE)
  if (missing(team)) {
   return(as_tibble(json_franchise$data))
    }
  else {
      filtered<-switch(type, 
                   num = (json_franchise$data) %>% filter(id == team),
                   char = (json_franchise$data) %>% filter(teamCommonName == team))
      }
  return(as_tibble(filtered))
}

franchise(26, "num")


franchise_team <- function(team, type) {
  team_url <- paste0("https://records.nhl.com/site/api/franchise-team-totals")
  get_franchise_team <- GET(team_url)
  txt_fran_team <- content(get_franchise_team, "text")
  json_fran_team <- fromJSON(txt_fran_team, flatten = TRUE)
  if (missing(team)) {
   return(as_tibble(json_fran_team$data))
    }
  else {filtered <- switch(type, 
                           num = (json_fran_team$data) %>% filter(franchiseId == team),
                           char = (json_fran_team$data) %>% filter(teamName == team))
  }
  return(as_tibble(filtered))
  }

franchise_team(15, "num")

season_record <- function(ID){
  season_url <- paste0("https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId=", ID)
  get_season <- GET(season_url)
  txt_season <- content(get_season, "text")
  json_season <- fromJSON(txt_season, flatten = TRUE)
  return(as_tibble(json_season$data))
}

season_record(ID= 24)  

goalie_record <- function(ID){
  goalie_url <- paste0("https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId=", ID)
  get_goalie <- GET(goalie_url)
  txt_goalie <- content(get_goalie, "text")
  json_goalie <- fromJSON(txt_goalie, flatten = TRUE)
  return(as_tibble(json_goalie$data))
}

goalie_record(24)

skater_record <- function(ID){
  skater_url <- paste0("https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId=", ID)
  get_skater <- GET(skater_url)
  txt_skater <- content(get_skater, "text")
  json_skater <- fromJSON(txt_skater, flatten = TRUE)
  return(as_tibble(json_skater$data))
}

skater_record(24)

hockey_stats <- function(endpoint, team){
  stats_url <- paste0("https://statsapi.web.nhl.com/api/v1/teams", endpoint)
  get_stats <- GET(stats_url)
  txt_stats <- content(get_stats, "text")
  json_stats <- fromJSON(txt_stats, flatten = TRUE)
  names <- json_stats$teams$name
  if (endpoint == "?expand=team.roster") {
  filtered <- list(json_stats$teams[["roster.roster"]], .name_repair = "unique")
  }
  else if (endpoint == "?expand=person.names"){
    filtered <- as_tibble(data.frame(json_stats$teams))
  }
  else if (endpoint == "?expand=team.schedule.next") {
    filtered <- list(json_stats$teams[["nextGameSchedule.dates"]])
  }
  else if (endpoint == "?expand=team.schedule.previous") {
    filtered <- list(json_stats$teams[["previousGameSchedule.dates"]])
  }
  else if (endpoint == "?expand=team.stats"){
    filtered <- list(json_stats$teams$teamStats[["splits"]])
  }
  else if (endpoint == "?expand=team.roster&season=20142015") {
    filtered <- list(json_stats$teams[["roster.roster"]])
  }
  else if (endpoint == "?teamId=4,5,29") {
    filtered <- list(json_stats$teams)
  }
  else {
    filtered <- list(json_stats$teams)
  }
  return(list(filtered))
  }


hockey_stats("?stats=statsSingleSeasonPlayoffs")
```

# Wrapper
```{r wrapper, warning= FALSE}
wrapper <- function(endpoint, ...) {
  if (endpoint == "/franchise" ) {
    franchise()
  }
  else if (endpoint == "/franchise-team-totals") {
    franchise_team()
  }
  else if (endpoint == "/franchise-season-records?cayenneExp=franchiseId=") {
    season_record(...)
  }
  else if (endpoint == "/franchise-goalie-records?cayenneExp=franchiseId=") {
    goalie_record(...)
  }
  else if (endpoint == "/franchise-skater-records?cayenneExp=franchiseId=") {
    skater_record(...)
  }
  else {
    hockey_stats()
  }
}

wrapper("/franchise")
wrapper("/franchise-season-records?cayenneExp=franchiseId=", 24)
```

# Analytics
## Data Join
This is a join to look at franchise information and all goalie information. This data specifically pulls information for the Washington Capitals.  
``` {r data analysis, warning= FALSE}
#Join data 
right_join(franchise_team(), goalie_record(24), by= c("teamName" = "franchiseName"))
```
## New Variables
I created new variables to look at the average number of games played in a season for goalies and skaters as well as the average goals per season for skaters.  
```{r new var, warning= FALSE}
#2 new vars
goalie_record_newVar <- goalie_record(24) %>% mutate(avg_games= gamesPlayed/seasons)
skater_record_newVar <- skater_record(24) %>% mutate(avg_games = gamesPlayed/seasons) %>% mutate(goals_season = round(goals/seasons, 0))
```
## Contingency Tables
Below looks at the number of goals scored by each position, the number of goals by number of seasons played, and the number of shootout wins per team. 
For number of goals per position, you would expect the sum of the forward players to be more than that of the defensemen. For the number of goals by seasons played, you would expect the number of goals to increase as seasons increased. I wouldn't expect much of a pattern for shootouts, but you can see which teams are better at shootouts.
```{r contingency tables, warning= FALSE}
#contingency tables
table(skater_record_newVar$positionCode, skater_record_newVar$goals)
table(skater_record_newVar$seasons, skater_record_newVar$goals_season)
table(franchise_team()$teamName, franchise_team()$shootoutWins)
```
## Numerical Summaries
I created numeric summries by grouping on categorical variables and making calculations for quantitative data.  
The first looks at the total regular and post season wins by team. The second looks at the average number of goals scored by position for the Carolina Hurricanes.  
```{r numerical summaries, warning= FALSE}
#numerical summaries for quantitative data
franchise_team() %>% group_by(teamName) %>% summarise(regularandPostWins = sum(wins))
skater_record(26) %>% group_by(positionCode) %>% summarise(avgPts = mean(goals))
```

## Bar Plot
Below is a bar plot for the Nashville Predators. It looks at the total number of players that have played each position in the history of the franchise. Defensemen is the highest as you typically have multiple players on the ice playing this position at one time.  
```{r bar plot, warning= FALSE}
#Bar Plot
skater_position <- ggplot(data = skater_record(34), aes(x= positionCode))
skater_position+geom_bar(fill = "#ffb81c", col = "#041e42", size = 2) + labs(x="Position", title= "Number of players for each position for the Nashville Predators") +scale_fill_manual(values=("#ffb81c")) + scale_x_discrete(labels = c("Center", "Defenseman", "Left Wing", "Right Wing"))
```

## Histogram
This is a histogram of number of goals scored by each player of the Washington Capitals. It is high close to 0 because many players will only play for a team for a season, and many will not score a goal during this time. They can't all be Ovechkin.  
``` {r histograms, warning= FALSE}
#Histogram
goals_scored <- ggplot(data = skater_record(24), aes(x= goals))
goals_scored + geom_histogram(fill = "#C8102E", color = "#041E42", size = 2, binwidth = 50) + labs(x= "Goals", title = "Goals Scored by the Washington Capitals")
```

## Box Plot
The graph below shows a boxplot of total games played by position. There isn't really a pattern here, just that is is bounded at 0 because many players on a roster will not play in a season, or will only play a few games with the given franchise.  
The other graph shows the same data, but line charts for the mean number of games played. It is parsed out by players currently on the team and players not active on the team.  
```{r boxplots, warning= FALSE}
#Boxplot
skater_position2 <- ggplot(data = skater_record(32), aes(x=positionCode, y=gamesPlayed))
skater_position2 + geom_boxplot(fill = "#F47A38", col = "#00685E") + labs(x="Position", y= "Games Played", title = "Box Plots of Games Played by position for the Anaheim Ducks")
skater_position2 +  stat_summary(fun = mean, geom = "line", lwd= 1.5, aes(group = activePlayer, col= activePlayer)) + labs(c="Position", y= "Games Played", title = "Mean games played by position for the Anaheim Ducks by active player")
```

## Scatterplot
The last scatterplot shows average games per season by total wins for goalies. There appears to be a positive trend, that the more seasons a goalie plays, the more wins they will have. This is also parsed out between active and non active players. There appears to be about 3 active goalies for the Capitals.  
```{r scatterplot, warning= FALSE}
#scatterplot
goalie_scat <- ggplot(data = goalie_record_newVar, aes(x=avg_games, y=wins, color = activePlayer))
goalie_scat + geom_point() + labs( x = "Average Games per Season", y = "Total Wins", title = "Total wins by Average games per Season for Washington Capitals Goalies")
````
